---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

### EpiInvert (version 0.3.1, December 2022)

<!-- badges: start -->
<!-- badges: end -->

Using variational techniques we address some epidemiological problems as the incidence curve decomposition or the estimation of the functional relationship between epidemiological indicators. We also propose a learning method for the short time forecast of the trend incidence curve. 

#### Vignettes of the main package functionalities 

* [`EpiInvert`](https://lalvarezmat.github.io/EpiInvert/EpiInvertVignette.html) : an incidence curve decomposition by inverting the renewal equation.

* [`EpiInvertForecast`](https://lalvarezmat.github.io/EpiInvert/EpiInvertForecastVignette.html) : a learning method for the short time forecast of the trend incidence curve. 

* [`EpiIndicators`](https://lalvarezmat.github.io/EpiInvert/EpiIndicatorsVignette.html) : estimation of the delay and ratio between epidemiological indicators. 

We also present in [Rt Comparison](https://lalvarezmat.github.io/EpiInvert/RtComparisonVignette.html) 
a comparative analysis of the methods : EpiInvert, 
[EpiEstim]( https://CRAN.R-project.org/package=EpiEstim), 
[Wallinga-Teunis](https://academic.oup.com/aje/article/160/6/509/79472)
and 
[EpiNow2]( https://CRAN.R-project.org/package=EpiNow2).

#### EpiInvert Installation

You can install the development version of EpiInvert from [GitHub](https://github.com/) with:

```{r setup, eval=FALSE}
 install.packages("devtools")
 devtools::install_github("lalvarezmat/EpiInvert")
```

#### Example
We attach some required packages
```{r, echo = TRUE,message = FALSE,warning=FALSE }
library(EpiInvert)
library(ggplot2)
library(dplyr)
library(grid)
```  

Loading data on COVID-19 daily incidence up to 2022-05-05 for 
[France](https://www.santepubliquefrance.fr/dossiers/coronavirus-covid-19/coronavirus-chiffres-cles-et-evolution-de-la-covid-19-en-france-et-dans-le-monde), 
[Germany](https://experience.arcgis.com/experience/478220a4c454480e823b17327b2bf1d4), 
[the USA](https://ourworldindata.org/coronavirus-source-data) 
and [the UK](https://ourworldindata.org/coronavirus-source-data): 
```{r}
data(incidence)
tail(incidence)
```

Loading some festive days for the same countries: 
```{r}
data(festives)
head(festives)
```

Executing EpiInvert using Germany data:
```{r,echo = TRUE, message = FALSE,results = FALSE}
res <- EpiInvert(incidence$DEU,"2022-05-05",festives$DEU)
```

Plotting the results: 
```{r fig1, fig.width = 5, fig.height = 8, fig.align = "center"}
EpiInvert_plot(res)

```

For a detailed description of EpiInvert outcomes see the [EpiInvert vignette](https://lalvarezmat.github.io/EpiInvert/EpiInvertVignette.html). 



