---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EpiInvert

<!-- badges: start -->
<!-- badges: end -->

EpiInvert estimates time varying epidemic reproduction numbers and restored 
incidence curves by inverting a renewal equation through a variational model as 
described in 
[PNAS, 2021](https://www.pnas.org/doi/10.1073/pnas.2105112118) and 
[Biology, 2022](https://www.mdpi.com/2079-7737/11/4/540). 
EpiInvert also corrects the administrative weekly bias in the daily registration 
of cases and the bias introduced by the festive days.  EpiInvert can manage 
daily incidence data and weekly aggregated incidence data where every 
week a single data is communicated with the accumulated incidence in the last 
7 days. 

## Vignettes

* [EpiInvert](https://ctim.ulpgc.es/covid19/EpiInvertVignette.html) :
A detailed description of EpiInvert R package functionalities.

* [Rt Comparison](https://ctim.ulpgc.es/covid19/RtComparison.html) : 
A comparative analysis of the methods : EpiInvert, 
[EpiEstim]( https://CRAN.R-project.org/package=EpiEstim), 
[Wallinga-Teunis](https://academic.oup.com/aje/article/160/6/509/79472)
and 
[EpiNow2]( https://CRAN.R-project.org/package=EpiNow2).


## EpiInvert Installation

You can install the development version of EpiInvert from [GitHub](https://github.com/) with:

```{r setup, eval=FALSE}
 install.packages("devtools")
 devtools::install_github("lalvarezmat/EpiInvert")
```

## Example
We attach some required packages
```{r, echo = TRUE,message = FALSE }
library(EpiInvert)
library(ggplot2)
library(dplyr)
library(grid)
```  

Loading data on COVID-19 daily incidence up to 2022-05-05 for 
[France](https://www.santepubliquefrance.fr/dossiers/coronavirus-covid-19/coronavirus-chiffres-cles-et-evolution-de-la-covid-19-en-france-et-dans-le-monde), 
[Germany](https://experience.arcgis.com/experience/478220a4c454480e823b17327b2bf1d4), 
[the USA](https://ourworldindata.org/coronavirus-source-data) 
and [the UK](https://ourworldindata.org/coronavirus-source-data): 
```{r}
data(incidence)
tail(incidence)
```

Loading some festive days for the same countries: 
```{r}
data(festives)
head(festives)
```

Executing EpiInvert using Germany data:
```{r,echo = TRUE, message = FALSE,results = FALSE}
res <- EpiInvert(incidence$DEU,"2022-05-05",festives$DEU)
```

Plotting the results: 
```{r fig1, fig.width = 5, fig.height = 8, fig.align = "center"}
EpiInvert_plot(res)

```

For a detailed description of EpiInvert outcomes see the [EpiInvert vignette](https://ctim.ulpgc.es/covid19/EpiInvertVignette.html). 



